---
title: "Getting Started with osfr"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with osfr}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

This vignette provides a quick tour of the `osfr` package. 

```{r message=FALSE}
library(osfr)
library(dplyr)

# external file for our example project
write.csv(mtcars, "mtcars.csv")
```


## Background 

[OSF][osf], developed by the [Center for Open Science (COS)][cos], is a free, open source, web application that provides a space for researchers to collaboratively store, manage, and share their research materials (e.g. data, code, protocols). Files can be kept private, shared with a specific group of researchers, or made publicly available with citable DOIs so researchers can get credit for their work.

`osfr` allows researchers to create and modify projects, components, and files through OSF's API.

## Working with projects/components

Most work on OSF is organized around ***projects***. Projects can contain *files*, groups of files in *directories*, and/or files in subsections of a project called ***components***. Collectively projects and components are called ***nodes***.

### Creating a new project/component

At the start of a new research project or study we'll want to set up a new private project on OSF to store all the files related to the project. We can add multiple nested components or directories to create whatever structure will suite our workflow. Different components can have different DOIs and different privacy settings, so we may want to put similar files in their own components to organize the project.

We'll first create a new *project* and store osfr's response in the `car_project` object. This object contains a specialized data frame based on the [tibble][] class that provides the new project's global unique identifier (GUID), title, and other metadata returned by the API. 

*Note: You'll need to be authenticated to create `nodes`. See the `osf_auth()` function documentation or the `auth` vignette for more information.*

```{r}
car_project <- osf_create_project("Motor Trend Car Road Tests")
car_project
```

To view the project directly on OSF use `osf_open(car_project)`.

After the project is created we'll create two nested *components*, one for data and one for analysis scripts. We can take advantage of osfr's [pipe][magrittr]-compatibility to arrange multiple operations into a single sequence.

```{r}
scripts_comp <- car_project %>%
  osf_create_component("Data") %>%
  osf_create_component("Analysis Scripts")
```

Because each operation in the sequence above applies to the previous object, we've accidentally created a project with the `analysis scripts` component nested within the `data` component, which is not what we wanted. We can delete the analysis scripts component and recreate it under the top-level project.

```{r}
osf_rm(scripts_comp, check = FALSE)

car_project %>% 
  osf_create_component(title = "Analysis Scripts")
```

We can verify the project is now structured the way we wanted by listing the components we have under the main project.

```{r}
car_components <- osf_ls_nodes(car_project)
car_components
```

Remember, new projects are always private by default. You can change this by opening the project on OSF and [making it public][osf-privacy], and [even generate a DOI][osf-doi] so others can cite your work. 

## Working With Files

In general, files can exist in projects, components, and/or directories on OSF. Files can be stored on *OSF's Storage* or in another service that is connected to an OSF project (e.g. GitHub, Dropbox, or Google Drive). However, `osfr` currently only supports interacting with files on OSF Storage.

### Uploading/Updating files

To add our data file to the *Data* component within our project, we'll need to pass an `osf_tbl` containing only that component to the `osf_upload()` function, so the file is uploaded to the correct part of the project.

*__Note:__ If an `osf_tbl` with multiple entities is passed to an osfr function, the default behavior is to use the entity in the first row and warn that all other entities are ignored.*

osfr is compatible with the dplyr package, so we can `filter()` for the specific component we want.

```{r}
library(dplyr)

# select the data component
data_component <- filter(car_components, name == "Data")

# upload our file
mtcars_file <- osf_upload(data_component, path = "mtcars.csv")
mtcars_file
```

The `mtcars_file` object now contains an `osf_tbl` with 1 row corresponding to the version of that file stored on OSF. We can also use `osf_open()` to examine the file directly on OSF, which provides rendered views for a wide variety of file formats.

OSF also version controls files, which means that if we were to make a change to the `mtcars.csv` file (perhaps we collected more data) we could automatically create a 2nd version of the dataset by using the `osf_upload()` function and setting `conflicts = "overwrite"`. As long as the file we're uploading has the same file name as a file in that particular section of the project, a new version will be created.

### Downloading files

We can download files from any public or private node that we have access to and can identify files to download in two different ways.

1. If we know where the file is located, but don't remember it's GUID, we can use the `osf_ls_files` function to list all the files within a specified node and then pipe the results to `osf_download()`.

    ```{r message=FALSE}
    data_component %>% 
      osf_ls_files() %>%
      osf_download(conflicts = "overwrite")
    ```

2. For a public file that we may have seen referenced in a published article, we may already have the GUID, and so can retrieve the file directly before downloading it. Here, we'll download Daniel Laken's helpful spreadsheet for calculating effect sizes (<https://osf.io/vbdah/>).

    ```{r eval=FALSE}
    excel_file <- osf_retrieve_file("vbdah")
    osf_download(excel_file)
    ```

```{r cleanup, include=FALSE}
osf_rm(car_project, recursive = TRUE, check = FALSE)
unlink("mtcars.csv")
```


## More Information

For more information on OSF and `osfr` check out:

* [OSF][osf]
* [OSF API Documentation][osf-api]
* [OSF Support](https://osf.io/support/)
* [osfr GitHub Repository](https://github.com/CenterForOpenScience/osfr)

<!-- links -->
[osf]: https://osf.io
[cos]: https://cos.io
[osf-api]: https://developer.osf.io
[magrittr]: https://magrittr.tidyverse.org
[tibble]: https://tibble.tidyverse.org
[osf-privacy]: https://help.osf.io/hc/en-us/articles/360018981414-control-your-privacy-settings
[osf-doi]: https://help.osf.io/hc/en-us/articles/360019931013-Create-DOIs
